---
title: "Capture-Mark-Resight Mabuias"
author: "Tati Micheletti"
date: "1 May 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage{color}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mabuyas

The present analysis focused on determining the density of mabuyas (*Trachylepis atlantica*) in different habitats on the island of Fernando de Noronha. The `R` package used for the present analysis is the `RMark` and the code is hosted in `github` (https://github.com/tati-micheletti/mabuyas). This `.rmd` file will guide you on how to reproduce the analysis. Make sure you also have program [Mark](http://www.phidot.org/software/mark/docs/book/) and [RStudio](www.rstudio.com) installed in your computer. Mark needs to be installed in the ProgramFiles folder (`C:\ProgramFiles\`).

### Getting your project
1. Go to [\textcolor{red}{github website}](https://github.com/tati-micheletti/mabuyas).
2. Click on the blue button that says `clone or download`, and click on the `download ZIP` to download the folder.
3. Now you are set to start working on your project, so open the file `mabuyas.Rproj` in `RStudio`. 
4. Now, on the lower-right pannel, click on the file mabuyasAnalysisDirectDownload.Rmd and you will open this source file.
5. Run all code chuncks below by pressing `Ctrl + Alt + R` or clicking in `Run` and `Run All` in the tab above.

### Description of analysis using `RMark` package in `R`
\
Analysis were performed using the [`Poisson-log normal mark-resight model`](http://www.phidot.org/software/mark/docs/book/pdf/chap18.pdf).

1. Install and load the libraries:

```{r load libraries}
suppressWarnings(library("RMark"))
suppressWarnings(library("xlsx"))
suppressWarnings(library("reproducible"))
```

If you don't have the libraries above, install and run them again:

```{r install libs, eval = FALSE}
install.packages("RMark")
install.packages("xlsx")
install.packages("reproducible")
```


## Create your data.frame and variables

First of all, to analyse your data, you need to bring it to `R`.

2. Create a data.frame in R from `excel` to be analysed by Mark:
```{r data}
directory <- getwd()
ch <- read.xlsx(file = file.path(getwd(), "data/dataMabuya.xlsx"), 
                       header = TRUE, 
                       sheetName = "ch", 
                       as.data.frame = TRUE)
ch$ch <- as.character(ch$ch) # Convert factor to character

covars <- read.xlsx(file = file.path(getwd(), "data/dataMabuya.xlsx"), 
                       header = TRUE, 
                       sheetName = "covars", 
                       as.data.frame = TRUE)
```

3. Create the variables you will need for the model(s) in form of a vector or matrix from your datasets:
```{r variables}
nocc <- 6 # Number of occasions
groupsNames <- unique(covars$colonies) # Names of groups
nGroups <- length(groupsNames) # Number of groups

unmarkedSeen <- matrix(covars[, "unmarkedSeen"], 
                       nrow = nGroups, 
                       ncol = nocc, 
                       byrow = TRUE) # counts of unmarked individuals seen
markedUnidentified <- matrix(covars[, "markedUnidentified"], 
                             nrow = nGroups, 
                             ncol = nocc, 
                             byrow = TRUE) # counts of marked individuals unidentified
knownMarks <- matrix(covars[, "knownMarks"], 
                     nrow = nGroups, 
                     ncol = nocc, 
                     byrow = TRUE) # counts of known marked individuals
effort <- matrix(covars[, "effort"], 
                 nrow = nGroups, 
                 ncol = nocc, 
                 byrow = TRUE) # effort of observation in each colony at each time
```

Now you have all the basic data to run your models. But first, you need the models.

## Defining your model(s)

To define your models, you have to think of which variables will influence your parameters. For example: different effort in each occasion will influence the probability of observing individuals ($\alpha$), so $\alpha$ could vary in function of effort or even, both effort and colony ('colonies') depending if you think that it is harder or easier to observe the individuals in different places. We can also come up with extra categories related to disturbance level (i.e. capim-açu is less disturbed than boldro), for example. Here I ran only one model so far, but you should later work on more models and we can develop a comparison test. :)

Basically you have to give a formula to each variable. for example: the basic formula `formula = ~1` means that the parameter that this formula is assigned to is constant. We can also make this parameter vary with `effort` (`formula = ~ effort`) or even to `effort` and `colonies` as suggested before (`formula = ~ effort + colonies`). Within a model formula you can also set parameters to fixed values. For example, we know that no mabuyas died within a week of work, so we can set survival (`Phi`) to `1`. We do that by using the constant formula and adding the argument `fixed = 1` (`formula = ~ 1, fixed = 1`). It is also important that all models are in a `list()`. All parameters are descibed below in section *6. Make a data design*.  

4. Defining the model
```{r models}
alpha.effort <- list(formula = ~ effort)
Phi <- list(formula = ~ 1, fixed = 1)
U <- list(formula = ~ colonies) 
constant <- list(formula = ~ 1) # Here I create a generic constant 
                                # model to be used to terms that I need
zero <- list(formula = ~ 1, fixed = 0) # Here I create a generic zero 
                                       # model to be used to terms that I need
```

## Process your data to make it ready for the analysis

To have the data ready **with** the covariates you want to consider (i.e. `effort`), you need to go through 2 steps before running your model:  
\textcolor{blue}{1. Process your data}: use the function `process.data()`  
\textcolor{blue}{2. Make a data design}: use the function `make.design.data()`  

5. Process your data
```{r process data}
mabuya.process <- process.data(ch,
                               model = "PoissonMR", 
                               time.intervals = rep(1, times = (nocc-1)),
                               groups = "colonies",
                               counts = list("Unmarked Seen" = unmarkedSeen,
                                             "Marked Unidentified" = markedUnidentified,
                                             "Known Marks" = knownMarks))
mabuya.process
```

Now you can see a list that summarizes your data. You can see the variables by typing `names(mabuya.process)`:  

```{r names mabuya.process}
names(mabuya.process)
```

And you can access these elements by typing `mabuya.process$ELEMENT_NAME`.

6. Make a data design:

Now you need to make a data design that will identify the parameters in your data that are constant, and the variable ones for each model term, as well as set your covariates where they should be placed in:  

```{r data design}
mabuya.ddl <- make.design.data(mabuya.process)
```

Using the `names(mabuya.ddl)` you can check now your parameters. 
```{r names ddl}
names(mabuya.ddl)
```

The *\textcolor{red}{Poisson-log normal mark-resight model}*, has have *four* main parameters:

\textcolor{blue}{U[j]} (`U`) = number of unmarked individuals in the population during primary interval j.  
\textcolor{blue}{$\alpha[j]$} (`alpha`) = intercept (on log scale) for mean resighting rate during primary interval j. If there is no individual heterogeneity ($\sigma[j]$ = 0), once back-transformed from the log scale the real parameter estimate can be interpreted as the mean resighting rate for the entire population.  
\textcolor{blue}{$\varphi[j]$} (`Phi`) = apparent survival between primary intervals j and j + 1.  
\textcolor{blue}{$\sigma^2[j]$} (`sigma`) = individual heterogeneity level (on the log scale) during primary interval j, i.e., the additional variance due to a random individual heterogeneity effect with mean zero.  

Now we can check one of the parameters: `alpha`:
```{r alpha}
mabuya.ddl$alpha
```

Now you can include the covariates (such as effort and/or disturbance level) with the following code:
```{r add covariates}
mabuya.ddl$alpha$effort <- as.numeric(apply(X = mabuya.ddl$alpha, 1, function(x){
    col <- x[["group"]]
    occ <- as.numeric(x[["time"]])
    eff <- covars[covars$colonies==col&covars$occasion==occ, "effort"]
  })
)
```

And now we check again to see how this changed our `data.frame`:
```{r alpha2}
mabuya.ddl$alpha
```

Here, we added effort to alpha (which is the mean resighting rate during primary interval), as more effort increases the chances of observing an individual and the effort varied on occasions and colonies.

Now we are ready to run the model(s).

## Running the analysis

7. Model coding
```{r run_model}
setwd(dir = file.path(getwd(), "outputs")) # Changing directory for outputs
mabuya.model <- mark(data = mabuya.process,      # Here we use the processed data 
                     ddl = mabuya.ddl,           # Here we use the designed data 
                     model.parameters = list(    # Here are the parameters's   
                             alpha = alpha.effort, # models set on step 4.
                             Phi = Phi,
                             U = U,
                             sigma = zero,
                             GammaDoublePrime = constant,
                             GammaPrime = constant
                     )
                 )
setwd(dir = directory) # Changing directory back to original
```

## Interpreting the results

Now, it is important to notice that the `U` the model outputs is `NOT` the  `population size`. The last, as well as other parameters can, however be derived:  

*Derived Parameters*

\textcolor{blue}{$\lambda[j]$} = overall mean resighting rate for primary occasion j. This is a parameter derived as a function of $\alpha[j]$, $\sigma^2[j]$ and and E[j]. Note that when $\sigma[j]$ = 0 and E[j] = 0, then the real parameter estimate for $\alpha[j]$ is identical to the derived parameter estimate for $\lambda[j]$.

\textcolor{blue}{N[j]} = U[j] + n[j], or the total population size during primary occasion j. This is a derived parameter, because MARK actually estimates U[j] in the model. n[j] is the marked population.
  
8. Calculating derived parameters

RMark automatically calculates the `population size` for each group (and occasion when we do a long term study with longer interval time).
```{r derived}
mabuya.model$results$derived$`N Population Size`
```

You will notice, however, that it doesn't output the location's names. We can, however retrieve it from the original data, and clean it up (repetitions):

```{r derived Final}
populationSize <- cbind(covars$colonies, mabuya.model$results$derived$`N Population Size`) %>%
  unique()
populationSize <-  populationSize[populationSize$se > 0,]
```

Now we can also calculate the densities among the areas (and at some point compare them statistically):
```{r density}
populationSize$areaSize <- c(36.7, 44.92, 7.7,
                             3.5, 15, 52, 48, 30)
density <- populationSize$estimate/populationSize$areaSize
populationSize <- cbind(populationSize, density)
colnames(populationSize) <- c("location", "estimate", "standard.error",
                           "lowerCI", "upperCI", "areaSize", "density")
```

And here we see the final result of the population size:
```{r final, echo = FALSE}
knitr::kable(populationSize, digits = 2, align = "c", format = "pandoc", row.names = FALSE)
```
