---
title: "Capture-Mark-Resight Mabuias"
author: "Tati Micheletti"
date: "1 May 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mabuyas

The present analysis focused on determining the density of mabuyas (*Trachylepis atlantica*) in different habitats on the island of Fernando de Noronha. The `R` package used for the present analysis is the `RMark` and the code is hosted in `github`. This `.rmd` file will guide you on how to reproduce the analysis. Make sure you also have program [Mark](http://www.phidot.org/software/mark/docs/book/) and [RStudio](www.rstudio.com) installed in your computer. Mark needs to be installed in the ProgramFiles folder (`C:\ProgramFiles\`).

### Setting a github account and creating your repository for the project
The first step is to set a github account so you can download the code and we can work on it collaboratively:
1. Go to [github website](www.github.com), and sign up for an account.
2. In the search bar, look for me `tati-micheletti`. On the left bar my name will be shown under `users`. Click to follow me.
3. Click on my profile and go to my `repositories`. You will see one that is called `mabuyas`. Click on it.
4. Now you can see all the files (this guide included!) that are there. There is a little button on the top right that says `fork`. Click on it. This will copy my repository to a repository on your account.
5. Now we need to install [git](https://git-scm.com/download/win) and [gitkraken](www.gitkraken.com) in your computer. This will allow us to work on your project in a collaborative way.
6. Once  both are installed, create an account in gitkraken (you can actually link this to your github account).
7. Now go back to your github account and click in repositories and in the `mabuya` one. Now we need to get all those files to your computer. Click on the green button that says `clone or download`, and click on the little icon on the right of the link to copy it.
8. Go back to your gitkraken and click on `File` and `Clone Repo`. Put the url you just copied, define a folder for the project on `Where to clone` and click on `Clone the repo!`.
9. Now you are set to start working on your project.

### Analysis using `Mark` program:
1. Create the file `.inp` with observations (created for you in folder data): `secondAnalysis.inp`. This file contains information on capture history, groups, number of individuals marked, number of unmarked individuals that were observed, and number of individuals that are marked and you can identify individually. The file also has infomation on the effort spent in each occasion. 
2. 

### Alternative analysis using `RMark` package in `R` (NOT FINISHED!)

Analysis were performed using the [`Poisson-log normal mark-resight model`](http://www.phidot.org/software/mark/docs/book/pdf/chap18.pdf).

1. Install and load the libraries:

```{r load libraries}
install.packages("RMark")
library("RMark")
```

## Create your data.frame

2. Create a data.frame in R from `excel` to be analysed by Mark:
```{r data}
mabuya.df <- 
```


## Defining parameters and model

3. Define parameter and the model you want to use
```{r load data}
mabuya.process <- process.data(mabuya.df,model="Poisson-log normal mark-resight model", time.intervals=c(1,1,1,1,1), groups="colonies")
```



