---
title: "Capture-Mark-Resight non-spatial wirh RMark"
author: "Tati Micheletti"
date: "1 May 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mabuyas

The present analysis focused on determining the density of mabuyas (*Trachylepis atlantica*) in different habitats on the island of Fernando de Noronha. The `R` package used for the present analysis is the `secr`. Analysis were performed based on [secr documentation for R](https://www.otago.ac.nz/density/pdfs/secr-markresight.pdf).

```{r load libraries}
install.packages("secr")
require("secr")
```

## Defining the model

According to Table 1 (secr 3.1), the model to be used is Model 1 (Capture-mark-resight (1,. . . ,0), with All ID, without correction). First, we convert the original data into a `capthist` format for the analysis. The first thing is to define the `markocc`(mark occasions). When marking occasion occured, we give it value 1. If occasion only relates to resighting, we give it a 0. Information on data input can be found [here](https://www.otago.ac.nz/density/pdfs/secr-datainput.pdf)

```{r load data}

mark.occ <- 1 # Occasion of marking
sight.occ <- 5 # Number of sighting occasions
markocc(traps) <- c(mark.occ, sight.occ)


```
